Individual and Differential Harm in Redistricting

### 2022-08-27 paper/harm.Rmd
add clarity to the limitations ~ some are cute but useless

A voter is harmed only if they are not represented by the candidate of their choice under $q$ but they would have been represented by their candidate of choice under the counterfactual plan $\tilde q$.
Notice the fundamental asymmetry in this definition: $h_i(v,q,\tilde q)\neq h_i(v,\tilde q,q)$.
Exchanging the role of the enacted plan and the counterfactual will not simply flip the direction of our conclusion.
In general, $q$ will harm some individuals while $\tilde q$, and $\tilde q$ will harm a different set of individuals.
A symmetric version of $h_i(v,q,\tilde q)$ would need to combine both of these groups: those harmed and those who gain by electing their chosen candidate under $q$ but not under the counterfactual $\tilde q$.
In practice, voters, courts, or politicians may find different balances in preventing harm and maximizing gains.

Our definition of harm is limited in two key regards.
First, it considers only individuals who vote: people who abstain from voting have a harm of zero.
A natural extension of this definition is to separate individual candidate preference from vote choice. 
The latter defines election winners, but the former provides a more-broad harm measure.
Naturally, this approach requires auxiliary data to implement, such as a large-scale survey, as candidate preferences for non-voters are not directly inferred from election returns.

>>>>

A voter is harmed only if they are not represented by the candidate of their choice under $q$ but they would have been represented by their candidate of choice under the counterfactual plan $\tilde q$.
Notice the fundamental asymmetry in this definition: $h_i(v,q,\tilde q)\neq h_i(v,\tilde q,q)$.
Exchanging the role of the enacted plan and the counterfactual will not simply flip the direction of our conclusion.
In general, $q$ harms some individuals while $\tilde q$, and $\tilde q$ harms a distinct, though not necessarily disjoint, set of individuals.
A symmetric version of $h_i(v,q,\tilde q)$ would need to combine both of these groups: those harmed and those who gain by electing their chosen candidate under $q$ but not under the counterfactual $\tilde q$.
In practice, voters, courts, or politicians may find different balances in preventing harm and maximizing gains.

Our definition of harm is limited in two notable regards.
First, it considers only individuals who vote: people who abstain from voting have a harm of zero.
Empirically, this changes nothing, but limits the measure in a theoretical sense.
People who do not voter express no formal preference, even if they have informal preferences.
The natural theoretical extension of this definition is to then separate individual preferences from vote choice. 
Naturally, extending this approach requires auxiliary data to implement, such as an absurdly large-scale survey, as candidate preferences for non-voters are not directly inferred from election returns.
As surveys are typically benchmarked against electoral outcomes, we do not endorse this type of extension.
